Wavefield Reconstruction Inversion: an example

### 2020-03-28 repro/papers/wri1d/paper.tex
near final - need DSO ref

The main result of this paper is that in one simple case, in which all
of the necessary computations can be carried out by hand, this hope is
not realized: minimization of the WRI objective is just as likely to
be trapped in a spurious local minimizer as is the FWI objective.  The
context of this conclusion is a simple transmission inverse problem
for the 1D acoustic wave system, which models pulse transmission along
a 1D continuum from a source point to a receiver point. The pulses
used in this thought experiment are short, so the main information
content of the data is the time of transit from source to
receiver. The predominant information about the material model, in
this case reduced to the wave velocity (a scalar function of
position), is just this transit time, so I constrain both the target
wave velocity $c_*$ generating the data and the trial wave velocity
$c$ to be constant, that is, independent of position along the 1D
continuum. I introduce a family of inverse problems, depending on a
parameter $\lambda$ playing the role of wavelength. For sufficiently small
$\lambda$, it is possible to show explicitly that the FWI problem
possesses local minimizers far from the global minimizer at the target
$c_*$, and that initiating a local iterative optimization, such as
steepest descent or Newton's method, at a distance from $c_*$ bounded
below by a multiple of the ``wavelength'' $\lambda$ will result in
convergence to these spurious local minima. That is, FWI 
behaves in exactly the manner described in much of the literature on
this topic. However, an analysis of WRI applied to the same context
yields the same result: spurious local minima exist for sufficiently
small $\lambda$, and will be found by local optimization unless the
starting point is within $O(\lambda)$ of the target. That is, WRI
behaves in a manner qualitatively indistinguishable from FWI. In
particular, its ability to allow good fit to data for small $\alpha$
does not safeguard it from failure to converge globally to a ``good''
local minimum. In fact, the $\alpha \rightarrow 0$ limit of the WRI
objective is well-defined (after scaling by $1/alpha^2$ and also behaves in
qualitatively the same way as the FWI objective. So the apparent
ability to maintain better data fit via reduction of $\alpha$ does not
lead to global behaviour asymptotically more amenable to local optimization.

@@

Recall that the object of this study is the global behaviour of
objective functions for velocity estimation: in this context, that
means the behaviour for $c$ far from $c_*$. Define
\begin{equation}
  \label{eqn:defL}
  L = \frac{2 c_{\rm max}^2}{|z_s-z_r|}.
\end{equation}
Then if $|c_*-c| > L\lambda$,
\[
  \left|\frac{|z_s-z_r|}{c}-\frac{|z_s-z_r|}{c_*}\right|
  =\frac{|c-c_*||z_s-z_r|}{cc_*} 
\]
\[
\ge \frac{|c-c_*||z_s-z_r|}{c_{\rm max}^2} > L\lambda
\frac{|z_s-z_r|}{c_{\rm max}^2} = 2\lambda.
\]
That is, the infima of the supports of $S_p[c]\wl, \dl=S_p[c_*]\wl$
are further apart than the lengths of these supports. In that case,
necessarily
\[
  \mbox{supp }S_p[c]\wl \cap \mbox{supp }S_p[c_*]\wl = \emptyset.
\]
Therefore $S_p[c]\wl$ and $S_p[c_*]\wl$ are orthogonal in $L^2[0,T]$,
and
\begin{equation}
  \label{eqn:fwiscrewed}
  J_{\rm FWI,\lambda}[c] =
  \frac{1}{2}\left(\frac{1}{4c^2}+\frac{1}{4c_*^2}\right)\|w_1\|^2.
\end{equation}
Amongst other consequences, one immediately deduces from the
expression \ref{eqn:fwiscrewed} the non-convexity result:

@@

As noted in the text, the point source $w(t)\delta(z-z_s)$ is not a
member of the domain of the simulation operator $S[c]$, as it is not
square-integrable. The object of this appendix is to construct a square-integral right-hand side in the
system \ref{eqn:awe1d} for which the pressure field $p$
is the same as that of the weak solution to the point source problem \ref{eqn:awe1dptsrc}
constructed in the last section, near the receiver point $z=z_r$, and
to exhibit this square-integrable replacement for the point source as
the image of the point source wavelet under a bounded extension map,
as in equation \ref{eqn:ext}.

>>>>

The main result of this paper is that in one simple case, in which all
of the necessary computations can be carried out by hand, this hope is
not realized: minimization of the WRI objective is just as likely to
be trapped in a spurious local minimizer as is the FWI objective.  The
context of this conclusion is a simple transmission inverse problem
for the 1D acoustic wave system, which models pulse transmission along
a 1D continuum from a source point to a receiver point. The pulses
used in this thought experiment are short, so the main information
content of the data is the time of transit from source to
receiver. The predominant information about the material model, in
this case reduced to the wave velocity (a scalar function of
position), is just this transit time, so I constrain both the target
wave velocity $c_*$ generating the data and the trial wave velocity
$c$ to be constant, that is, independent of position along the 1D
continuum. I introduce a family of inverse problems, depending on a
parameter $\lambda$ playing the role of wavelength. For sufficiently small
$\lambda$, it is possible to show explicitly that the FWI problem
possesses local minimizers far from the global minimizer at the target
$c_*$, and that initiating a local iterative optimization, such as
steepest descent or Newton's method, at a distance from $c_*$ bounded
below by a multiple of the ``wavelength'' $\lambda$ will result in
convergence to these spurious local minima. That is, FWI 
behaves in exactly the manner described in much of the literature on
this topic. However, an analysis of WRI applied to the same context
yields the same result: spurious local minima exist for sufficiently
small $\lambda$, and will be found by local optimization unless the
starting point is within $O(\lambda)$ of the target. That is, WRI
behaves in a manner qualitatively indistinguishable from FWI. In
particular, its ability to allow good fit to data for small $\alpha$
does not safeguard it from failure to converge globally to a ``good''
local minimum. In fact, the $\alpha \rightarrow 0$ limit of the WRI
objective is well-defined (after scaling by $1/\alpha^2$) and also behaves in
qualitatively the same way as the FWI objective. So the apparent
ability to maintain better data fit via reduction of $\alpha$ does not
lead to global behaviour asymptotically more amenable to local optimization.

@@

Recall that the object of this study is the global behaviour of
objective functions for velocity estimation: in this context, that
means the behaviour for $c$ far from $c_*$. Define
\begin{equation}
  \label{eqn:defL}
  L = \frac{2 c_{\rm max}^2}{|z_s-z_r|}.
\end{equation}
Then if $|c_*-c| > L\lambda$,
\[
  \left|\frac{|z_s-z_r|}{c}-\frac{|z_s-z_r|}{c_*}\right|
  =\frac{|c-c_*||z_s-z_r|}{cc_*} 
\]
\[
\ge \frac{|c-c_*||z_s-z_r|}{c_{\rm max}^2} > L\lambda
\frac{|z_s-z_r|}{c_{\rm max}^2} = 2\lambda.
\]
That is, the infima of the supports of $S_p[c]\wl, \dl=S_p[c_*]\wl$
are further apart than the lengths of these supports. In that case,
necessarily
\[
  \mbox{supp }S_p[c]\wl \cap \mbox{supp }S_p[c_*]\wl = \emptyset.
\]
Therefore $S_p[c]\wl$ and $S_p[c_*]\wl$ are orthogonal in $L^2[0,T]$,
and
\begin{equation}
  \label{eqn:fwiscrewed}
   J_{\rm FWI}[c,\dl,\wl]  =
  \frac{1}{2}\left(\frac{1}{4c^2}+\frac{1}{4c_*^2}\right)\|w_1\|^2.
\end{equation}
Amongst other consequences, one immediately deduces from the
expression \ref{eqn:fwiscrewed} the non-convexity result:

@@

As noted in the text, the point source $w(t)\delta(z-z_s)$ is not a
member of the domain of the simulation operator $S[c]$, as it is not
square-integrable. The object of this appendix is to construct a square-integrable right-hand side in the
system \ref{eqn:awe1d} for which the pressure field $p$
is the same as that of the weak solution to the point source problem \ref{eqn:awe1dptsrc}
constructed in the last section, near the receiver point $z=z_r$, and
to exhibit this square-integrable replacement for the point source as
the image of the point source wavelet under a bounded extension map,
as in equation \ref{eqn:ext}.
